---
title: "Choosing number of clusters in GoM model"
author: "Joyce Hsiao"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, message=F)
```

---

## Background and summar

The GoM (Grades of Membership) model computes log Bayes factor for comparing model fitting with k clusters versus null single cluster model. The log Bayes factor is corrected for the number of parameters fitted in each model. In clustering analysis, one typically chooses k that has the largest Bayes factor. However, it is worth remembering that in practice, it is unlikely to get a true k, and results from different k may complement each other rather than competing with each other. 

In the current analysis, we evaluated the data with k = 2 to 20. We found that k=2 produces the largest Bayes factor.

---

## Analysis

```{r}
library(knitr)
library(kableExtra)
library(dplyr)
library(metagenomeSeq)
library(CountClust)
```

Read in filtered data.

```{r}
MRobj <- readRDS("../data/nasal_filtered_normed_batchcorrected.rds")
counts <- MRcounts(MRobj,norm=FALSE,log=FALSE)

MRobj
```

Fitting k=2 to 10.

```{r, eval = F}
fits <- FitGoM(t(counts), K=c(2:10), tol=0.00001)
saveRDS(fits, file = "../data/gomfits.rds")
```

Extract sample memberships.

```{r}
fits <- readRDS(file = "../data/gomfits.rds")

clust_fit_membership <- lapply(fits, function(xx) {
     apply(xx$omega, 1, which.max)
  })

for (i in 1:length(clust_fit_membership)) {
  print(table(clust_fit_membership[[i]]))
}
```

---

## Log Bayes Facctor

```{r, echo = FALSE}
bf_species <- do.call(rbind, lapply(1:length(fits), function(i) {
  data.frame(no_cluster = c(2:10)[i], BF = fits[[i]]$BF)
}) )

bf_species_seq <- 
  data.frame(no_cluster = bf_species$no_cluster[-1],
             BF_change = (bf_species$BF[-1] - bf_species$BF[-length(bf_species$BF)])/bf_species$BF[-1] )
```


```{r, fig.width=7, fig.height=3}
par(mfrow=c(1,2))
plot(bf_species, type = "b", xlab = "Number of clusters",
     ylab = "log Bayes Factor (H1 vs single cluster)", main = "Species")
plot(bf_species_seq, type = "b", xlab = "Number of clusters",
     ylab = "Percent change in log Bayes Factor (k vs k-1)", main = "Species")
```



---

## Session information

```{r, echo = FALSE}
sessionInfo()
```

