---
title: "Compute variance explained under different K"
author: "Joyce Hsiao"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, message=F)
```

## Background and summary

We compare fitting under differet k and computed for every feature, sample variance in abundance that is explained by sample cluster membership. 

Results showed that k=2 gives the largest variance explained (median of sample variances is higher than the other k).

---

## Analysis

```{r}
library(ggplot2)
library(metagenomeSeq)
```

Load data

```{r}
MRobj <- readRDS("../data/nasal_filtered_normed_batchcorrected.rds")
counts <- MRcounts(MRobj,norm=FALSE,log=FALSE)
```

Load previously computed results.

```{r}
fits <- readRDS(file = "../output/chossing-number-of-clusters.Rmd/gomfits.rds")
```

Extract sampel cluster membership.

```{r}
membership <- lapply(fits, function(xx) {
     apply(xx$omega, 1, which.max)
  })
```

Compute variance explained.

```{r, eval = FALSE}
varprop_species <- lapply(1:length(membership), function(index) {
  which_sample <- which(colnames(counts) %in% names(membership[[index]]))
  counts_which <- counts[, which_sample]
  var <- sapply(1:nrow(counts_which), function(index2) {
            res <- lm(log2(counts_which[index2,]+2) ~ membership[[index]])
            sumOfsquares <- anova(res)[[2]]
            sumOfsquares[1]/sum(sumOfsquares) })
  return(var)
})
names(varprop_species) <- names(membership)
saveRDS(varprop_species, file = "../output/cluster-variance-explained.Rmd/varprop.rds")
```

Plot variance explained

```{r}
varprop <- readRDS("../output/cluster-variance-explained.Rmd/varprop.rds")
varprop_long <- data.frame(
  prop = do.call(c, varprop),
  cluster = as.factor(rep(2:20, each = length(varprop[[1]]))))

ggplot(varprop_long, aes(x = cluster, y = prop,
                            col = cluster), size = .2) +
  geom_violin(trim = FALSE, size=.5) +
  geom_boxplot(width=0.1, size = .6) +
  labs(x = "Number of clusters", y = "Proportion variance explained")

varprop_long$cluster <- as.numeric(as.character(varprop_long$cluster))
ggplot(subset(varprop_long, cluster  < 6), aes(x = as.factor(cluster), y = prop,
                            col = as.factor(cluster)), size = .2) +
  geom_violin(trim = FALSE, size=.5) +
  geom_boxplot(width=0.1, size = .6) +
  labs(x = "Number of clusters", y = "Proportion variance explained")

for (i in 1:4) {
  cat(c(2:5)[i], "clusters \n")
  print(summary(varprop[[i]]))
}
```


---

## Session information

```{r, echo = F}
sessionInfo()
```

