---
title: "Applying GoM to HMP data"
author: "Joyce Hsiao"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
#knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, message=F)
```

---

## Summary

* Apply hierarchical clustering to cluster samples
* Try different distance metrics
    + Euclidean on logged and normed counts
    + Center log ratio transform
    + Bray-Curtis on sample proportions
    

---

## Euclidean on logged and normed counts

On logged and normed counts. Complete and average distance give the same results. 

```{r}
library(metagenomeSeq)
MRobj <- readRDS("../data/nasal_filtered_normed_batchcorrected.rds")
counts_logged <- MRcounts(MRobj,norm=T,log=T)
colnames(counts_logged) <- colnames(MRobj)

dist_mat <- dist(t(counts_logged), method="euclidean")
hclust_avg <- hclust(dist_mat, method = 'complete')
cut_avg_1 <- cutree(hclust_avg, k = 2)
table(cut_avg_1)
plot(hclust_avg)
rect.hclust(hclust_avg , k = 2, border = c(2,4))

#saveRDS(cut_avg_1, file = "output/cluster-alternative-methods.Rmd/cut_res_1.rds")
```

Total tax count and cluster membership

```{r}
plot(cut_avg_1, colSums(MRcounts(MRobj,norm=T,log=F)))
t.test(colSums(MRcounts(MRobj,norm=T,log=F)) ~ cut_avg_1)
```







## Center log ratio transform

On normed counts. 

```{r}
library(metagenomeSeq)
MRobj <- readRDS("../data/nasal_filtered_normed_batchcorrected.rds")
counts <- MRcounts(MRobj,norm=T,log=F)
colnames(counts) <- colnames(MRobj)

library(compositions)
counts_clr <- t(counts)

# use cutree to determine sample cluster membership
dist_mat <- dist(counts_clr, method="euclidean")
hclust_avg <- hclust(dist_mat, method = 'complete')
cut_avg_2 <- cutree(hclust_avg, k = 2)
table(cut_avg_2)
plot(hclust_avg)
rect.hclust(hclust_avg , k = 2, border = c(2,4))

#saveRDS(cut_avg_2, file = "output/cluster-alternative-methods.Rmd/cut_res_2.rds")
```





## Bray-Curtis on sample proportions


```{r}
library(metagenomeSeq)
MRobj <- readRDS("../data/nasal_filtered_normed_batchcorrected.rds")
counts <- MRcounts(MRobj,norm=T,log=F)
colnames(counts) <- colnames(MRobj)

library(vegan)
mat_prop <- t(t(counts)/colSums(counts))
beta_bray <- vegdist(t(mat_prop), method = "bray", upper=T)

# use cutree to determine sample cluster membership
#dist_mat <- dist(t(counts_clr), method="euclidean")
hclust_avg <- hclust(beta_bray, method = 'complete')
cut_avg_3 <- cutree(hclust_avg, k = 2)
table(cut_avg_3)
plot(hclust_avg)
rect.hclust(hclust_avg , k = 2, border = c(2,4))

#saveRDS(cut_avg_3, file = "output/cluster-alternative-methods.Rmd/cut_res_3.rds")
```


##  Output hierarchical clustereing results.

```{r, eval=T}
cut_avg_1 <- readRDS(file = "../output/cluster-alternative-methods.Rmd/cut_res_1.rds")
cut_avg_2 <- readRDS(file = "../output/cluster-alternative-methods.Rmd/cut_res_2.rds")
cut_avg_3 <- readRDS(file = "../output/cluster-alternative-methods.Rmd/cut_res_3.rds")

all.equal(names(cut_avg_1), names(cut_avg_2))
all.equal(names(cut_avg_1), names(cut_avg_3))

df <- data.frame(sample=names(cut_avg_1),
                 method1=cut_avg_1,
                 method2=cut_avg_2,
                 method3=cut_avg_3)

# write.csv(df,
#             file= "../output-manuscript/supp-clusterig.csv",
#             quote=F, row.names=F)

table(df$method1)
table(df$method2)
table(df$method3)

table(df$method1, df$method3)
table(df$method1, df$method3)

obj <- readRDS("../data/nasal_GOM.rds")
gom <- pData(obj)$GOM
names(gom) <- rownames(pData(obj))

table(gom, df$method1)
table(gom, df$method2)
table(gom, df$method3)
```


---

## Session information

```{r}
sessionInfo()
```

