---
title: "Investigate clusering probability in relation to PCA results"
author: "Joyce Hsiao"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, message=F)
```

## Background and summary


---

## Analysis

```{r}
library(ggplot2)
library(metagenomeSeq)
library(matrixStats)
```

Load PCA

```{r}
obj <- readRDS("../data/nasal_GOM.rds")
```

Load previously computed results.

```{r}
fits <- readRDS(file = "../data/gomfits.rds")
```

Extract sampel cluster membership.

```{r}
membership <- lapply(fits, function(xx) {
     apply(xx$omega, 1, which.max)
  })
```

Get PCA results

```{r}
pcaRES <- function(obj,tran=TRUE,comp=1:2,norm=TRUE,log=TRUE,usePCA=TRUE,useDist=FALSE,distfun=stats::dist,dist.method="euclidian",n=NULL,...){
  mat = returnAppropriateObj(obj,norm,log)
  if(useDist==FALSE & usePCA==FALSE) stop("Classical MDS requires distances")
  if(is.null(n)) n = min(nrow(mat),1000)

  otusToKeep <- which(rowSums(mat)>0)
  otuVars<-rowSds(mat[otusToKeep,])
  otuIndices<-otusToKeep[order(otuVars,decreasing=TRUE)[seq_len(n)]]
  mat <- mat[otuIndices,]

  if(tran==TRUE){
    mat = t(mat)
  }
  if(useDist==TRUE){
    d <- distfun(mat,method=dist.method)
  } else{ d = mat }

  if(usePCA==FALSE){
    ord = cmdscale(d,k = max(comp))
    xl = paste("MDS component:",comp[1])
    yl = paste("MDS component:",comp[2])
  } else{
    pcaRes <- prcomp(d)
    ord <- pcaRes$x
    vars <- pcaRes$sdev^2
    vars <- round(vars/sum(vars),5)*100
    xl <- sprintf("PCA %s: %.2f%% variance",colnames(ord)[comp[1]], vars[comp[1]])
    yl <- sprintf("PCA %s: %.2f%% variance",colnames(ord)[comp[2]], vars[comp[2]])
  }
  return(pcaRes)
}

obj <- obj[,order(pData(obj)$GOM)]

pcares <- pcaRES(obj,pch=21,bg=pData(obj)$GOM,main="")
```

save to output folder. 

```{r}
saveRDS(pcares, "../output/explore-cluster-probability.Rmd/pcares.rds")
```


## Results

7 outliers on PCA plot. All of them below to the second cluster.


```{r}
all.equal(rownames(pcares$x), colnames(obj))
pdata <- pData(obj)
pdata$PC1 <- pcares$x[,1]
pdata$PC2 <- pcares$x[,2]
pdata$pc_outlier <- pdata$PC1 > 20 & pdata$PC2 < -20

all.equal(rownames(pdata), rownames(fits[[1]]$omega))
omega <- fits[[1]]$omega
omega <- omega[match(rownames(pdata),rownames(omega)),]
all.equal(rownames(pdata), rownames(omega))

pdata$gom_prob_group1 <- omega[,1]
pdata$gom_prob_group2 <- omega[,2]
```

* 7 PCA outliers, all belong to cluster 2, their cluster probability are not borderline.

```{r}
table(pdata$pc_outlier)

table(pdata$pc_outlier, pdata$GOM)

cbind(pdata$gom_prob_group1[which(pdata$pc_outlier==1)], pdata$gom_prob_group2[which(pdata$pc_outlier==1)])

```

* Are their borderline probablities at all?

```{r}
summary(pdata$gom_prob_group2[pdata$GOM==2 & pdata$pc_outlier==0])
summary(pdata$gom_prob_group2[pdata$GOM==2 & pdata$pc_outlier==1])
```

The non-outliers in Group 2 are more similar to Group 1. 

```{r}
library(CountClust)
newgroup <- pdata$GOM
newgroup[which(pdata$pc_outlier==1 & pdata$GOM=="2")] <- "2.outlier"
annotation <- data.frame(
  sample_id = paste0("X", c(1:NROW(omega))),
  tissue_label = factor(newgroup,
                        levels = c("1", "2", "2.outlier") ) ) 
rownames(omega) <- annotation$sample_id;

StructureGGplot(omega = omega,
                annotation = annotation,
                palette = RColorBrewer::brewer.pal(3, "Accent"),
                yaxis_label = "Group",
                order_sample = TRUE,
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 7,
                                 axis_label_face = "bold"))
```

Save outlier info.

```{r, eval = F}
write.table(data.frame(rownames(pdata),newgroup)[order(newgroup, decreasing = T),],
            col.names = F, row.names = F, quote = F,
            file = "../output/explore-cluster-probability.Rmd/outlier.list")
```





---

## Session information

```{r, echo = F}
sessionInfo()
```

