---
title: "Grades of membership model: choose random seed"
author: "Joyce Hsiao"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, message=F)
```

---

## Summary

Motivation: when running `CountClut`, sample cluster assignment may differ between different random seeds, despite similar cluster characteristics in feature abundancce. This issue is not unique to `CountClust`. In model-based unsupervised clustering methods, the parameter space is usually complex, and as a result, difficult to find global maximum of the model likelihood. 

I run `CountClust` under 100 different random seeds and then compare Bayes Factors across different random seeds and sample cluster membership. I will use the random seed that generates the largest Bayes Factor in all subsequent `CountClust` analysis. 

This document shows the analysis steps for when K=2. Results show that for when K=2, across 100 random seeds `409` produced the largest BF and is thus chosen as our random seed. This chosen random seed (best seed) is then used in our final results (https://jhsiao999.github.io/nasalmicrobiome/step3_gradesOfMembership.html). 

The fitting results for K=2 to 5 under the same 100 random seeds are saved in `output/gom-seeds-k-2.rds`, `output/gom-seeds-k-3.rds`, `output/gom-seeds-k-4.rds`, and `output/gom-seeds-k-5.rds`.



---

## Package

```{r}
library(knitr)
library(kableExtra)
library(dplyr)
library(metagenomeSeq)
library(CountClust)
library(parallel)
```

## Loading data

Read in filtered data.

```{r}
MRobj = readRDS("../data/nasal_filtered_normed_batchcorrected.rds")
MRobj
```


---

## K=2 

100 random seeds

Apply `CountClust` to fit grades of membership model. 

```{r, eval = T}
counts <- MRcounts(MRobj,norm=FALSE,log=FALSE)

# get 100 random seeds
source("../code/print-prime.R")
seeds <- prime(543)
```

```{r, eval = F}
res <- vector("list", length(seeds))

for (i in 1:100) {
  set.seed(seeds[i])
  res[[i]] <- FitGoM(t(counts), K=c(2), tol=1e-6)[[1]]
}

saveRDS(res, file = "../data/res.rds")
```


```{r, eval = T}
res <- readRDS(file="../data/res.rds")
```

Bayes factor differs between the seeds.

```{r}
summary(do.call(c,lapply(res, "[[", 4)))
```

Choose the seed with maximum BF: `r seeds[19]`

```{r}
which.max(lapply(res, "[[", 4))
```

```{r, eval=F}
saveRDS(res[[19]], file = "../output/gom-seeds.Rmd/res_best_seed_67.rds")
```

$~$

Results using the seed with the highest BF.

```{r, eval=T}
res_best <- readRDS(file = "../output/gom-seeds.Rmd/res_best_seed_67.rds")
```

```{r}
cluster <- apply(res_best$omega, 1, which.max)

table(cluster)
kable(res_best$omega)
```


---

## K=3 to 5

Code in `code/gom-seeds.R`.

```{r}
clust_2 <- readRDS("../output/gom-seeds-k-2.rds")
clust_3 <- readRDS("../output/gom-seeds-k-3.rds")
clust_4 <- readRDS("../output/gom-seeds-k-4.rds")
clust_5 <- readRDS("../output/gom-seeds-k-5.rds")
```

Identify the best seed.

K=2: 67; K=3: 67, K=4: 59; K=5: 307.

```{r}
seeds[which.max(lapply(clust_2, "[[", 4))]
seeds[which.max(lapply(clust_3, "[[", 4))]
seeds[which.max(lapply(clust_4, "[[", 4))]
seeds[which.max(lapply(clust_5, "[[", 4))]
```

save the best seeds.

```{r, eval =F}
saveRDS(clust_2[which.max(lapply(clust_2, "[[", 4))],
        "../output/gom-k2-best-seed-67.rds")
saveRDS(clust_3[which.max(lapply(clust_3, "[[", 4))],
        "../output/gom-k3-best-seed-67.rds")
saveRDS(clust_4[which.max(lapply(clust_4, "[[", 4))],
        "../output/gom-k4-best-seed-59.rds")
saveRDS(clust_5[which.max(lapply(clust_5, "[[", 4))],
        "../output/gom-k5-best-seed-307.rds")
```

---

## Session Info

```{r, echo = F}
sessionInfo()
```
