---
title: "Grades of membership model"
author: "JNP"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, message=F)
```

---

## Summary

Load data post filtering, normalization and batch correction (`nasal_filtered_normed_batchcorrected.rds`).

Apply grades of membership model using `CountClust`.

$~$

---

## Package

```{r}
library(knitr)
library(kableExtra)
library(dplyr)
library(metagenomeSeq)
library(CountClust)
```

## Loading data

Read in filtered data.

```{r}
MRobj = readRDS("../data/nasal_filtered_normed_batchcorrected.rds")
MRobj
```

$~$

---

## Clustering 

Apply `CountClust` to fit grades of membership model. 

```{r, eval = F}
set.seed(17)
counts_clust <- FitGoM(t(counts), K=c(2), tol=0.00001)
#saveRDS(counts_clust, file = "../data/count-clust.rds")
```

Load previous computed count_clust results.

```{r}
counts_clust <- readRDS(file = "../data/count-clust.rds")

cct <- counts_clust[[1]]$omega
rowMaxs = sapply(seq(nrow(cct)),function(i){
    seq(along=cct[i,])[cct[i,]==max(cct[i,])]
  })
rowMaxs = unlist(rowMaxs)
uniqueGroups = unique(unlist(rowMaxs))
groupIds = split(seq(rowMaxs),factor(rowMaxs))

cl = rep("1",sum(sapply(groupIds,length))); cl[groupIds[[2]]] = 2

table(cl)
cbind(groupIDs = cl,round(cct,3)) %>% kable %>%
  kable_styling()
```

$~$

---

## Preserve data with GOM

Next we save the dataset with the sample clustering labels

```{r, eval =F}
MRobj2 = MRobj
assayData(MRobj2)[["counts"]] = counts
featureData(MRobj2) = AnnotatedDataFrame(data.frame(Species=rownames(counts)))
pData(MRobj)$GOM = cl
pData(MRobj2)$GOM = cl

save(MRobj2, MRobj, file="../data/nasal_GOM.rdata")
```


$~$


---

## Session Info

```{r, echo = F}
sessionInfo()
```
